### 2017-06-01 R/oregon.Rmd
edits to Oregon Markdown

```{r misc2, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
activeTransportFile <- "~/ITHIM/inst/activeTransportTime.csv"
GBDFile <- "~/ITHIM/inst/gbd.csv"
```

@@

```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.baseline <- createITHIM(activeTransportFile = activeTransportFile,
                              GBDFile = GBDFile)
```

Sample Files
* [activeTransportTime.csv](https://github.com/syounkin/ITHIM/blob/master/inst/activeTransportTime.csv)
* [gbd.csv](https://github.com/syounkin/ITHIM/blob/master/inst/gbd.csv)

@@

Suppose we wanted to create a scenario in which the mean cycling time
and mean walking time are doubled.  We need only to change two means
in the model and nothing else.  To do so we could avoid creating input
files by simply updating the baseline object with a few new
parameters.

First we need to know the name of the parameters we want to update.
Currently the list of parameters resides in the help page for the
ITHIM class (`help("ITHIM")`).  Alternatively one can use
`getParameterNames`.

@@

Now we need to retrieve the walking and cycling means from the
baseline model.  To do so we use the method `getMeans`.  To find a
list of available methods use `getMethodsITHIM` with no arguments.

```{r misc23, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getMethodsITHIM()
meanVec <- getMeans(ITHIM.baseline)
meanVec
meanWalkingTime <- meanVec$walk; meanCyclingTime <- meanVec$cycle
scenarioParameters <- list(muwt = 2*meanWalkingTime, muct = 2*meanCyclingTime)
scenarioParameters
```

Now we use the list of parameters and the baseline ITHIM object to
create the scenario ITHIM object.

@@

Let's inspect it to be sure.

@@

Once the baseline and scenario ITHIM objects are defined, we may
compute the change in burden due to physical activity as follows.

@@

We use the NAtional Household Transportation Survey data stored as a _TravelSurvey_ R object to estimate mean walk and cycle times.

@@

* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/master/inst/roadInjuries.csv)

>>>>

Sample Files
* [activeTransportTime.csv](https://ithim.wiscweb.wisc.edu/wp-content/uploads/sites/110/2017/06/activeTransportTime.csv)
* [gbd.csv](https://ithim.wiscweb.wisc.edu/wp-content/uploads/sites/110/2017/06/gbd.csv)

@@

Please update the path and names in the code below to match the csv files you wish to use.

```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.baseline <- createITHIM(activeTransportFile = "~/Downloads/activeTransportTime.csv",
                              GBDFile = "~/Downloads/gbd.csv")
```

@@

Suppose we wanted to create a scenario in which the mean cycling time
and mean walking time are doubled.  We need only to change two means
in the model and nothing else.  To do so we could avoid creating input
files by simply updating the baseline object with a few new
parameters.
(note: this is the overall mean for the whole population. From this, _ITHIM_ will create means for each age & gender strata)

First we need to know the name of the parameters we want to update.
Currently the list of parameters resides in the help page for the
ITHIM class (`help("ITHIM")`).  Alternatively one can use the method
`getParameterNames`.

@@

Now we need to retrieve the walking and cycling means from the
baseline model.  To do so we use the method `getMeans`, which returns a vector of means.  To find a
list of available methods use `getMethodsITHIM` with no arguments.

```{r misc23, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
getMethodsITHIM()
```

We'll store the vector of means in a variable called _meanVec_ and then view that vector.
```{r misc322, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
meanVec <- getMeans(ITHIM.baseline)
meanVec
```

We can also pull out the mode-specific means from that vector and assign them to their own variables.

```{r misc3322, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
meanWalkingTime <- meanVec$walk
meanCyclingTime <- meanVec$cycle
```

Next we'll create a list of the parameters we wish to change in our scenario. Recall, we're interested in a scenario that has double the mean walking time and double the mean cycling time of the baseline scenario.

```{r misc3242, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
scenarioParameters <- list(muwt = 2*meanWalkingTime, muct = 2*meanCyclingTime)
scenarioParameters
```

_ITHIM_'s comparative risk assessment framework uses two models to describe the potential health impacts for small differences between them. In our case the two models differ in the active travel times of the population.  Above we created the baseline model, and now we use the list of changed parameters to `update` the baseline ITHIM object. This method creates a copy of the baseline ITHIM object with the changes we'd like to see in the new ITHIM object. We call this new object the scenario.

@@

Let's inspect the scenario ITHIM object to be sure it includes the changes we expect.

@@

Once the baseline and scenario ITHIM objects are defined, we may
compute the change in burden due to physical activity as follows. We provide a few examples to give you an idea of options for diseases and burden.

@@

We use the National Household Transportation Survey (NHTS) data stored as a _TravelSurvey_ R object to estimate mean walk and cycle times.

@@

* [roadInjuries.csv](https://ithim.wiscweb.wisc.edu/wp-content/uploads/sites/110/2017/06/roadInjuries.csv)
